Stack Safety as a Security Property

### 2020-07-08 main.tex
Overloading hookrightarrow

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer to throughout the paper
(\cref{sec:running-example}), a description of our threat model
(\cref{sec:threat}) and some technical preliminaries
(\cref{sec:prelim}).

\section{Running Example}
\label{sec:running-example}

\leo{We've said this literally in the previous sentence. What should this intro
  be in the new reorg?}
%We begin with a running example that will serve both to demonstrate several
%``stack-unsafe'' behaviors that our stack safety property should prohibit
%and to ground the technical presentation in later sections.

@@

\section{Threat Model}

\leo{Needs high level signposting, transitions}

@@

@@

% 
%\leo{TODO: Add citations of mitigation techniques}
%\rb{And for those techniques, explain how they reflect their motivating
%examples. What does each technique do for those examples? How are they
%connected to this work?}
% 
%\paragraph*{Bad Write}
% 
%\[
%44 : [\rsp - 5] \leftarrow r_4
%%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
%\]
% 
%Instruction 44 currently is
%$ [\rsp - 2] \leftarrow r_4$,
%%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
%with the
%effect of storing the contents of $r_4$ as the result of {\tt h}. If
%instead it wrote to location $\rsp - 5$ then it would be overwriting
%private data (the return address) belonging to {\tt g}.
% 
%\paragraph*{Bad Read}
% 
%\[
%30 : r_5  \leftarrow [\rsp - 6]
%%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
%\]
% 
%Instruction 30 currently is
%$ r_5 \leftarrow [\rsp - 4]$,
%%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
%with the
%effect of loading the argument of {\tt g} to $r_5$. If instead it
%loaded the contents of location $\rsp - 6$, it would obtain access
%to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
%outputs the contents of $r_5$, that value would immediately be made public.
% 
%\paragraph*{Bad Control Flow}
% 
%\[
%5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
%%5 : \jalr ~ r_4 ~ \rra ~ 0
%\]
%If a {\jalr} instruction was added to {\tt main} with its target being
%the middle of the code block of {\tt h}.\apt{describe result of this?}

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):
\rb{This $\hookrightarrow_P$ is often enough written just $\hookrightarrow$.
Overload everywhere?}

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_p \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow_P \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a function map $\codemap$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition{Stack Integrity} Let $\callmap$ be a call map and
$\codemap$ a function map. We say that a system satisfies {\em stack
  integrity} with respect to $\callmap$ and $\codemap$ if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow_P
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

@@

\definition{Stack Confidentiality}
Let $\callmap$ be a call map and $\codemap$ a function map. We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
and \(R = \ret{\mach_c}\):

$$\forall \MPT_{\mathit{call}} ~ \machT. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$

@@

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (call maps, function maps, return maps, and entry
maps): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

@@

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state pair
\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
is the first unmatched return such that for some $\pol_r$
\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
then $\mach_r$ corresponds to a return state for $\mach_c$:
\[\ret{\mach_c}\ \mach_r\]

@@

In such a system, all memory addresses and registers (including the program
counter) are enriched with an abstract metadata tag, which can point to the
representation of an arbitrary data structure. A micropolicy is thus defined by
a collection of rules to be checked as each instruction executes and together
define the transfer function of the security policy. Each rule inspects the tags
associated to the inputs of an instruction (PC, registers, memory) as well as
the instruction opcode itself, and produces one of two outcomes: either the
instruction is allowed to execute (and generates and assigns tags for the new PC
and the result of the operation), or the machine fail-stops with a policy
violation.

Efficient execution of these micropolicies relies on hardware implementations
such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
maps tag inputs to outputs and a mechanism to trap misses to software handlers
and thereby compute the transfer function of the software-defined policy.
Previous work has shown that a wide range of micropolicies can be defined and
enforced efficiently \rb{cite}.
%
\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
policies}

\paragraph*{A conservative policy}
%
The first micropolicy we present to enforce the definition of stack safety
developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
stack memory with ownership information associated to each stack frame (and to
each separate object inside that frames, a more finely grained access control
that we do not need to consider here), and registers holding stack pointers with
matching access permission information. Beyond those capabilities, the compiler
generates designated sequences of instructions with special tags that grant them
permission to manage the stack, say during calls and returns. \rb{Conceivably on
alloc operations, etc., which we have not yet mentioned.}

Our micropolicy works as follows. We assume for ease of exposition that no
arguments are passed on the stack \rb{Come back to this later?}. Initially, the
entire stack is tagged as unused with an $\tagNoDepth$ tag \rb{actually, the
``initial frame'' should be tagged with its proper depth}, and the {\PCname}
carries in its tag at all times the depth of the current activation, starting
with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
exactly at those points in the program where the contour of the stack safety
property changes:

@@

\item From caller to callee, when the machine executes an instruction tagged
  with $\tagCall$ as belonging to the call map. At this point, the current
  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
  the locations of the new frame is marked $\tagStackDepth{n + 1}$.

\item From callee back to caller, when execution reaches the corresponding
  return instruction for the call instruction, as defined by
  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
  Prior to this point, the locations of the old frame have been restored to
  $\tagNoDepth$.

@@

These sequences also enforce the $\SP$ discipline. When stack frames are of
fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
the header sequence, which is later checked after decrementing $\SP$ during the
exit sequence to authorize the return. The well-formedness of the blessed
sequences can be checked statically. Only programs whose sequences are
well-formed are protected by the micropolicy, although programs are still
protected (by fail-stopping) even if the intended sequences are missing. By
tracking the execution of the blessed sequences in the tag state of the
{\PCname} register, the micropolicy ensures that the entry sequences of calls
declared in the call map are executed correctly.

@@

\paragraph*{Testing the conservative policy}
%
This conservative policy can be shown to satisfy an eager \emph{testing
property}, which at each step of the computation performs integrity and
confidentiality tests for a stack of running variants, which are kept in
lockstep with the original machine.

\definition{Stack integrity test}
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition{Stack confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of relevant information on a stack whose
  elements represent variants of the machine by recording an initial machine
  state, an initial and current variant of the machine, a contour and a return
  map:
%
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
%
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}

\definition{Variant stack test}
%
  Our testing property extends the basic tests above to whole variant stacks.
  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
  $\mach$ safisfies the \emph{variant stack test} if for each variant element
  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a function map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth if:

%  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
%  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
%  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
%  testing property} is satisfied at call depth $d$, and write
%  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:

@@

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
    use of this vs the return case below.}

  \item If
    %
    (1) the machine takes a \emph{non-call, non-return step} $(\mach,
    \pol) \hookrightarrow_P (\mach', \pol', \obs)$ and the variant stack steps
    to $\vs \hookrightarrow \vs'$,
    %
    (2) the variant stack test $\mathit{\vs}_T ~
    (\mach, \pol) ~ \vs$ is satisfied, and
    %
    (3) the stack safety test
    $\mathit{stack}_T ~ d ~ \MPT ~ \vs'$ is coinductively satisfied at depth $d$
    and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
    \pol')$,
    %
    {\em then} we conclude $\mathit{stack}_T d ~ ((\mach, \pol) : \MPT) ~ vs$.
    \rb{omitting stack size condition}

>>>>

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer
to throughout the paper (\cref{sec:running-example}) and some
technical preliminaries (\cref{sec:prelim}).

\section{Stack {\underline{Un}}safety by Example}
\label{sec:running-example}

We begin with a running example that will serve both to demonstrate several
``stack-unsafe'' behaviors that our stack safety property should prohibit
and to ground the technical presentation in later sections.

\subsection{Running Example}

@@

\subsection{Attacker Model}

@@

\subsection{Unsafe Variations}
\sna{Variations here collides with state variations. I propose substitutions.}
\leo{I agree there is a collision, not sure about ``substitutions''}
\sna{Mutations, drawing from the randomized testing literature?}

The execution trace shown in \cref{fig:running-trace-b} is an example
of a ``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. In the rest of
this section we will go over examples of unsafe behaviors, before
formally capturing what stack-safe means formally and discussing
enforcement techniques from the literature in later sections. Throughout
the paper examples will be localized modifications of the example, by
substituting an instruction at a given PC, written:
\[\PCname :\mathit{inst}\]

\leo{TODO: Add citations of mitigation techniques}
\rb{And for those techniques, explain how they reflect their motivating
examples. What does each technique do for those examples? How are they
connected to this work?}

\paragraph*{Bad Write}

\[
44 : [\rsp - 5] \leftarrow r_4
%44 : ~ \sw ~ \rsp ~ r_4 ~ \negate 5
\]

Instruction 44 currently is
$ [\rsp - 2] \leftarrow r_4$,
%$\sw ~ \rsp ~ r_4 ~ \negate 2$,
with the
effect of storing the contents of $r_4$ as the result of {\tt h}. If
instead it wrote to location $\rsp - 5$ then it would be overwriting
private data (the return address) belonging to {\tt g}.

\paragraph*{Bad Read}

\[
30 : r_5  \leftarrow [\rsp - 6]
%30 : ~ \lw ~ r_5 ~ \rsp ~ \negate 6
\]

Instruction 30 currently is
$ r_5 \leftarrow [\rsp - 4]$,
%$\lw ~ r_5 ~ \rsp ~ \negate 4$,
with the
effect of loading the argument of {\tt g} to $r_5$. If instead it
loaded the contents of location $\rsp - 6$, it would obtain access
to the value of {\tt main}'s local variable {\tt x}. Worse, since the next instruction
outputs the contents of $r_5$, that value would immediately be made public.

\paragraph*{Bad Control Flow}

\[
5  : \rra \leftarrow \PCname + 1 ; ~ \PCname \leftarrow r_4
%5 : \jalr ~ r_4 ~ \rra ~ 0
\]
If a {\jalr} instruction was added to {\tt main} with its target being
the middle of the code block of {\tt h}.\apt{describe result of this?}

\subsection{Policies}

@@

@@

We similarly overload  \(\mpstatename \hookrightarrow \MPT\),
to relate a machine-policy state pair to the MP-trace
induced by \(\stepstopol{}\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):
%\rb{This $\hookrightarrow$ is often enough written just $\hookrightarrow$.
%Overload everywhere?}

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstoobs{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

@@

\definition{Transition} Given a trace $\MPT$, a call map $\callmap$ and
a code map $\codemap$, an $n$-{\em transition} is a pair of
machine-policy state pairs $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$
in the trace such that:
\begin{itemize}
\item one state pair steps to the other:
  $$\exists \obs. ~ (\mach_c, \pol_c) \stepstopol{\obs} (\mach_e, \pol_e),$$
\item $\mach_c$ is a valid call with $n$ arguments:
  $$\callmap(\mach_c(\PCname)) = n,$$
\item and $\mach_e$ is an entry point for $f$:
  $$\codemap(\mach_e(\PCname)) = f.$$
\end{itemize}

\definition{Stack Integrity} Let $\callmap$ be a call map and
$\codemap$ a code map. We say that a system satisfies {\em stack
  integrity} with respect to $\callmap$ and $\codemap$ if, for every
initial machine-policy state pair $(\mach_0, \pol_0)$ and its induced
trace $\MPT$, and every $n$-transition $(\mach_c, \pol_c)$ and
$(\mach_e, \pol_e)$ in $\MPT$, the prefix $\MPT_{pre}$ of $\MPT$ from
$(\mach_e, \pol_e)$ until the first return corresponding to $\mach_c$
(if any) satisfies trace integrity with respect to the contour at the
call:
%
$$\forall \MPT_{pre}. ~ \mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{pre} ~ | ~ \ret{\mach_c} \Rightarrow
\mathit{int}~\mathit{Cof}(\mach_c, n)~\MPT_{pre}$$

@@

\definition{Stack Confidentiality}
Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c, f, n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, f, n)\)
and \(R = \ret{\mach_c}\):

$$\forall \MPT_{\mathit{call}} ~ \machT. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow \machT ~ | ~ \ret{\mach_c} \Rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ (\ret{\mach_c}) ~ \MPT_{\mathit{call}} ~ \machT_{\mathit{call}}$$

@@

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (code maps, function maps, return maps, and entry
maps): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

@@

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state pair
\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
is the first unmatched return such that for some $\pol_r$
\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
then $\mach_r$ corresponds to a return state for $\mach_c$:
\[\ret{\mach_c}\ \mach_r\]

@@

In such a system, all memory addresses, registers and the program counter are
enriched with an abstract metadata tag, which can point to the representation of
an arbitrary data structure. A micropolicy is thus defined by a collection of
rules to be checked as each instruction executes and together define the
transfer function of the security policy. Each rule inspects the tags associated
to the inputs of an instruction (PC, registers, memory) as well as the
instruction opcode itself, and produces one of two outcomes: either the
instruction is allowed to execute (and tags for the new PC and its result are
produced and assigned to those), or the machine fail-stops with a policy
violation.

Efficient execution of these micropolicies relies on hardware implementations
such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
maps tag inputs to outputs and a mechanism to trap misses to software handlers
and thereby computes the transfer function of the software-defined policy.
Previous work has shown that a wide range of micropolicies can be defined and
enforced efficiently \rb{cite}.
%
\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
policies}

\paragraph*{A conservative policy}
%
The first micropolicy we present for eager stack safety is very similar to the
Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
tag the stack memory with ownership information associated to each stack frame
(and to each separate object inside that frames, a more finely grained access
control that we do not need to consider), and registers holding stack pointers
with matching access permission information. Beyond those capabilities, the
compiler generates designated sequences of instructions with special tags that
grant them permission to manage the stack.

Our micropolicy works as follows. We assume for ease of exposition that no
arguments are passed on the stack \rb{Come back to this later?}. Initially, the
entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
its tag at all times the depth of the current activation, starting with
$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
points in the program where the contour of the stack safety property changes:

@@

\item From caller to callee, when the machine executes an instruction tagged as
  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
  the new frame is marked $\tagStackDepth{n}$.

\item From callee back to caller, when execution reaches the instruction
  immediately following the original call instruction and the $\SP$ is the same
  as at the time of that call. At this point, the {\PCname} is decremented;
  previously, the locations of the old frame are restored to $\tagNoDepth$.

@@

These sequences also enforce the $\SP$ discipline. When stack frames are of
fixed size, it suffices to increment $\SP$ and give it with a special tag in the
header sequence, which is later checked after decrementing $\SP$ during the exit
sequence to authorize the return. The well-formedness of the blessed sequences
can be checked statically. Only programs whose sequences are well-formed are
protected by the micropolicy, although programs are still protected (by
fail-stopping) even if the intended sequences are missing. By tracking the
execution of the blessed sequences in the {\PCname} tag state, the micropolicy
ensures that the entry sequence is executed correctly a declared call.

@@

\paragraph*{Testing the conservative policy}
%
This conservative policy is can be shown to satisfy an eager \emph{testing
property}, which enforces lockstep and performs integrity and confidentiality
tests for a stack of running variants at each step.

\definition{Eager integrity test}
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{eager integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
    \]

\definition{Eager confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
  test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of execution information on a stack whose
  elements record an initial machine, an initial and current variant of the
  machine, a contour and a return map:
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
  with respect to it if:
  \begin{itemize}
  \item The current variant is indeed a variant of the current state.
  \item The stack is nonempty and the bottom return map is the false property.
  \item The current state is in the trace of every initial machine.
  \item The variant state is in the trace of its own initial variant.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item All contours have public registers.
  \end{itemize}

\definition{Eager testing property}
  Let $\callmap$ a call map and $\codemap$ a code map. Given a state
  $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$, we
  say that an MP-trace induced from $(\mach, \pol)$ satisfies the \emph{eager
  testing property} at a given call depth if:

@@

  \item If the machine fail-stops,
    $(\mach, \pol) \hookrightarrow (\mach, \pol, \tau)$,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
    depth up to the size of $\vs$.

  \item If the top variant of the stack (and only the top variant) returns,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
    equal to the size of the stack. \rb{Avoid confusion with return below}

  \item If the machine takes a step
    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
    that is neither a call nor a return, all
    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
    (these tests are required also for calls and returns, and omitted in those cases),
    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
    then the property is also satisfied for the extension of the trace \rb{notation}
    at depth $d$.

  \item If the machine takes a call step
    $(\mach, \pol) \hookrightarrow (\mach', \pol', \obs)$,
    the contour $\contour$ is updated based on the top of the current machine.
    For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    (and in addition to the general step conditions),
    the eager integrity test must hold for $\contour$ at the call,
    the eager confidentiality test must hold for all variants at the call,
    including the new variant that is added to the top of the stack,
    and the property recurses on an arbitrary variant at the new contour
    at depth $d$ with trace $\MPT$.
